Police Shootings and Racism in America

### 2020-10-16 police_modeling.Rmd
added to data section

```{r load-data}
all <- read.csv("allshootingsdata.csv")

philly <- read.csv("philly_shootings.csv")
cincinnati <- read.csv("cincinnati_shootings.csv")
```

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

>>>>

```{r load-data}
# Shootings
all <- read.csv("data/allshootingsdata.csv")
all <- all %>% 
  mutate(id = paste0("C", row_number()))

# Census Predictors
census <- read.csv("data/2013census.csv")[,-c(8,9)]
states <- lapply(census$department, function(x) str_extract_all(x,"\\([^()]+\\)")[[1]])
states <- substring(states, 2, nchar(states)-1)
census$state <- states

# Gun Law Predictors
gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
  filter(year == 2013)
gun_provisions <- gun_provisions %>% 
  mutate(total_pr = rowSums(gun_provisions[,3:136]),
         deesc_train = ifelse(state %in% c("California", "New Mexico", "Texas", "Oklahoma", "South Dakota", "Illinois", "Indiana", "Tennessee", "Washington", "Missouri", "Georgia", "Ohio", "Maryland", "Massachusetts", "Connecticut", "Maine"), 1, 0)) %>% 
  select(state, total_pr, deesc_train)
gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
```

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

### 2020-10-16 submissionii.Rmd
added to data section

### Data
|        As a basis for this study, I will be using the same dataset that VICE News used. This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects over seven years.” [1] The predictors that I will use are subject (victim) race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying. The response I will be modeling will be whether the shooting was fatal or not. I will be filtering out the cases in which fatality is unknown or the subject’s or the officer’s race or gender is unknown from the final dataset. Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information. I will carry out the same procedure in the cases in which there were more than one officers present at the shooting. Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.

@@

```{r load-data}
# Shootings
all <- read.csv("data/allshootingsdata.csv")
all <- all %>% 
  mutate(id = paste0("C", row_number()))

@@

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

@@

@@

```{r weapon-eda}
temp <- all %>% 
  group_by(Fatal, weapon) %>% 
  summarise(count = n())

ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings") 
```

@@

|        As I mentioned above, in order to draw comparisons between the demographics of those involved in police shootings versus non-police shootings, I will be using data collected by the Police Data Initiative on the shootings in two cities (Philadelphia and Cincinnati). I will only be using data from these two cities because obtaining data on both fatal and non-fatal shooting crime in America is extremely difficult, since non-fatal shootings are “generally lumped-in with other types of assault in national data collections.” [2] The Police Data Initiative also suggests using general shooting data from Rochester, New York. However, after looking into this dataset, I found that it did not specify whether each shooting was fatal or non-fatal. As a result, I have decided to leave Rochester out of my final dataset so that I can primarily focus on identifying trends in victim and officer demographics in the context of fatal versus non-fatal shootings. The main predictors I will use from the Philadelphia and Cincinnati datasets will be race and sex, both of which refer to the victims of the shootings. I will combine these selected variables from the two datasets into one final dataset. Once again, I will be using fatality of the shooting as the response variable.

```{r clean-cities}
cinn <- cincinnati %>% 
  select(RACE, SEX, TYPE) %>% 
  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
  mutate(city = "Cincinnati")

phil <- philly %>% 
  select(race, sex, fatal) %>% 
  rename(gender = sex) %>% 
  mutate(city = "Philadelphia")

comp <- rbind(cinn, phil)

comp <- comp %>%
  filter(race != "UNKNOWN") %>% 
  filter (race != "") %>% 
  filter(race != "UNK") %>% 
  filter(gender != "")

comp$race[comp$race == "WHITE"] <- "W"
comp$race[comp$race == "BLACK"] <- "B"
comp$race[comp$race == "OTHER"] <- "O"
comp$race[comp$race == "HISPANIC"] <- "H"

comp$gender[comp$gender == "MALE"] <- "M"
comp$gender[comp$gender == "FEMALE"] <- "F"

comp$fatal[comp$fatal == "0"] <- "N"
comp$fatal[comp$fatal == "1"] <- "F"
comp$fatal[comp$fatal == "NONFATAL"] <- "N"
comp$fatal[comp$fatal == "FATAL"] <- "F"
```

```{r fatal-eda1}
temp <- comp %>% 
  group_by(fatal) %>% 
  summarise(count = n())

ggplot(temp, aes(x=fatal, y=count)) + 
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
```

|        In this dataset, it is important to keep in mind that again, there is a significantly higher number of non-fatal shooting events than fatal shooting events.

```{r race-eda-cities}
temp <- comp %>% 
  group_by(fatal, race) %>% 
  summarise(count = n())

ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
```

```{r gender-eda-cities}
temp <- comp %>% 
  group_by(fatal, gender) %>% 
  summarise(count = n())

ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
```

|        Furthermore, it appears that Black is the race with the highest count and male is the gender with the highest count for victims of all shootings in these two cities, which is similar to what we discovered for the police shooting data. However, it does not seem that this trend in race and gender varies too much between fatal and non-fatal shootings.

@@

### Model 
```{r}

```

>>>>

@@

```{r load-data}
# Shootings
all <- read.csv("data/allshootingsdata.csv")

@@

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         id = paste0("C", row_number())) %>% 
  select(-weapon)

@@

### Data
|        As a basis for this study, I will be using the same dataset that VICE News used. This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including date, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not. Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), and Unknown (U). Gender was broken up into 3 categories: Male (M), Female (F), and Unknown (U). Additionally, there were cases in which multiple victims and/or multiple officers were present in the shooting. Each of these scenarios was still represented within one row of the dataset, but semi-colons were present in the respective victim and officer columns, separating each individual's information from each other.

@@

```{r weapon-eda, eval=F}
temp <- all %>% 
  group_by(Fatal, weapon) %>% 
  summarise(count = n())
ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings")
```

@@

@@

|        In the VICE dataset, I will be filtering out the cases in which fatality is unknown, or the subject’s or officer’s race or gender is unknown from the final dataset. Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information. I will carry out the same procedure in the cases in which there were more than one officers present at the shooting. Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.

### Model 
The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying. The response I will be modeling will be whether the shooting was fatal or not.
```{r}
fuck
```
